---
title: "Untitled"
author: "Xiaoyuan"
date: "2024-02-22"
output: html_document
---

# the following exercises are a test in disguise.

# what do you think about the coding?

# can you think of any improvements to the following?

```{r}
rm(list=ls()) # always clean up your environment! 
```

```{r}
# install.packages("tidyverse") #uncomment to install tidyverse if you haven't already

library(tidyverse)
```

## PROBLEM 1

# visualizing your data is important!

\# summary statistics can be highly misleading, and simply plotting the data can reveal a lot more! \# Lets look at the Anscombe’s Quartet data. There are four different data sets. \# Anscombe, F. J. (1973). "Graphs in Statistical Analysis". American Statistician. 27 (1): 17–21. <doi:10.1080/00031305.1973.10478966>. JSTOR 2682899.

```{r dataset setup}
anscombe_quartet <-  readRDS("../data/anscombe_quartet.rds")


# let's explore the dataset 
str(anscombe_quartet)

# what does the function str() do? 
# Display the structure of the dataset
```

```{r statistics summary}
# let's check some summary statistics:

result <- anscombe_quartet %>% 
  group_by(dataset) %>% 
  summarise(
    mean_x    = mean(x),
    mean_y    = mean(y),
    min_x     = min(x), 
    min_y     = min(y),
    max_x     = max(x),
    max_y     = max(y),
    crrltn    = cor(x, y)
  )
result

# what do the summary statistics tell us about the different datasets? 
# similar summary statistics (mean, min and max, correlation) for both x and y
```

```{r plot}
# let's plot the data with ggplot:

require(ggplot2)

ggplot(anscombe_quartet, aes(x=x,y=y)) +
  geom_point() + 
  geom_smooth(method = "lm",formula = "y ~ x") +
  facet_wrap(~dataset)

#save the plot

ggsave("../plot/anscombe_quartet.png", width = 20, height = 20, units = "cm")
```

```{r conclusion for question 1}
# what do the plots tell us about the different datasets? 
# they have notably different distributions and scatterplots

# describe the relationship between x and y in the different datasets.
# dataset_1: it shows a relatively strong positive linear relationship between x and y.
# dataset_2: y has a sooth curved relation with x.
# dataset_3: despite having an outlier, it also shows a strong positive linear relationship between x and y.
# dataset_4: Most of the points are clustered on x = 8, and an outlier is located at x = 19, which affects the regression pattern.

# would linear regression be an appropriate statistical model to analyse the x-y relationship in each dataset?
# No. Although the datasets are completely different, they show same regression coefficient and similar strong positive linear relationships.

# what conclusions can you draw for the plots and summary statistics? The fact that the summary statistics for the x and y variables in datasets are very similar does not mean that their distributions and scatterplots are also very similar. It is important to visualize data for a comprehensive understanding.
```

## PROBLEM 2

```{r data preparation}
# load in the datasaurus dataset
datasaurus_dozen <-  readRDS("../data/datasaurus_dozen.rds")

# explore the dataset 
str(datasaurus_dozen)

# how many rows and columns does the datasaurus_dozen file have? 
# there are 1426 rows and 3 columns in the dataset
```

```{r plot of whole dataset}
# plot the dataset 
# calculate the correlations and summary statistics for x and y in all datasets: 
result_2 <- datasaurus_dozen %>% 
  summarise(
    mean_x    = mean(x),
    mean_y    = mean(y),
    min_x     = min(x),
    min_y     = min(y),
    max_x     = max(x),
    max_y     = max(y),
    crrltn    = cor(x, y)
  )
result_2

ggplot(datasaurus_dozen, aes(x=x,y=y)) +
  geom_point() + 
  geom_smooth(method = "lm",formula = "y ~ x") +
  ggtitle("Datasaurus Dozen Scatterplot") +
  labs(x = "X-axis", y = "Y-axis") 
ggsave("../plot/datasaurus_dozen.png", width = 20, height = 20, units = "cm")
```

```{r plot of each dataset}
# Plot the relationships between x and y in each dataset including the line of best fit.
# save the plot 

ggplot(datasaurus_dozen, aes(x=x,y=y)) +
  geom_point() + 
  geom_smooth(method = "lm",formula = "y ~ x") +
  ggtitle("Datasaurus Dozen Scatterplot of different datasets") +
  labs(x = "X-axis", y = "Y-axis") +
  facet_wrap(~dataset)
ggsave("../plot/datasaurus_dozen_datasets.png", width = 20, height = 20, units = "cm")
```

```{r conclusion for question 2}
# what conclusions can you draw for the plots and summary statistics? 
# For the whole dataset, data points are randomly dispersed in a circle shape with a diameter equals to 100.
# For different datasets, they have very similar mean, max, min and correlation but different data distributions.
# conclusion: The fact that the summary statistics for the x and y variables in datasets are very similar does not mean that their distributions and scatterplots are also very similar. It is important to visualize data for a comprehensive understanding.
```
